---
title: "Basic data manipulation and visualization"
author: "Aishwarya"
date: "January 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("knitr")
#install.packages("markdown")
 #uninstall_tinytex()
library(knitr)
library(markdown)
 
```
                             


Let us write a function to subset a given  dataset

```{r }
library(ggplot2)
sub <- function(data, ...){
  arg <- list(...)
  a = data[0]
    
  for(i in arg){
    
    a <- cbind(a, data[i])
    }

return(a)
}
```

Testing it on mpg dataset
```{r cars}
sub(mpg, "drv" ,3, "cyl", 1, 4)
```


Now, writing a function to plot each column of dataset.If it’s a continuous variable (numeric), create a histogram. If it’s a categorical variable (character or factor),
create a bar plot.
```{r pressure, echo=TRUE}
plot <- function(data){
  for(i in names(data)){
    for(x in data[i]){
      if(is.numeric(x)){
        gh <- ggplot(data, mapping = aes(x = x)) + geom_histogram(bins = 10, fill ="yellow" , color ="green" ) + labs(x = i) 
       
        print(gh)
      }
      else{
        gb <- ggplot(data, mapping = aes(x = x)) + geom_bar(fill = "blue", color = "red") + labs(x=i) 
        print(gb)
      }
    }
  }
  
}

plot(mpg)
```





Use side-by-side boxplots to visualize the distribution of price for each level of color.
```{r}
ggplot(diamonds, mapping= aes(x = color, y = price)) + geom_boxplot(mapping= aes(fill = color)) + 
  ggtitle("Box Plot of Diamond Color Vs Price")

```
We notice that,
Worst diamonds have larger spread and less outliers compared to best diamonds.
People are paying high for better quality diamonds(G) rather than best(D).


Use side-by-side boxplots to visualize the distribution of carat for each level of color.

```{r}
ggplot(diamonds, mapping = aes(x = color, y = carat, fill = color)) + geom_boxplot() +
  ggtitle("Box Plot of Diamond Color Vs Carat")
```
             
Best diamonds has less weight whereas worst are more heavier.
From previous plot, people are paying more for Heavier and cheaper diamonds.                           
                                        
                                        
scatter plot of carat versus price, using either an additional aesthetic or faceting to visualize the
relationship between carat and price for each level of color.

```{r}

ggplot(data = diamonds,
       mapping = aes(x= carat, y= price)) +
  geom_point() + geom_smooth(method = loess) + facet_wrap(~color)
```
                                        
It can be said that most of the best diamonds has low weights and are sold at relatively lower prices than the better and worst diamonds. 
It is strange that none of the best quality diamonds weighs more than 3.5 carats and couldn't cross \$16000 whereas even worst diamonds with far lesser weights costs around \$18000.
                                        