---
title: "Boston-Housing Data"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(dplyr)
```

Fit a model that predicts per capita crime rate by town (crim) using only one predictor variable. Use plots
to justify your choice of predictor variable and the appropriateness of any transformations you use

```{r}
library(modelr)
library(readr)
library(dplyr)
library(ggplot2)
library(mlbench)

data(BostonHousing) 


data <- BostonHousing %>% as_tibble()

p1 <- ggplot(data) +geom_point(aes(log(crim), zn))
p2 <- ggplot(data) +geom_point(aes(log(crim),log(indus)))
p3 <- ggplot(data) +geom_boxplot(aes(y = log(crim), x = chas))
p4 <- ggplot(data) +geom_point(aes(log(crim), log(nox)))
p5 <- ggplot(data) +geom_point(aes(log(crim), log(rm)))
p6 <- ggplot(data) +geom_point(aes(log(crim),log( age)))
p7 <- ggplot(data) +geom_point(aes(log(crim), log(dis)))
p8 <- ggplot(data) +geom_point(aes(x =log(crim), y = log(rad)))
p9 <- ggplot(data) +geom_point(aes(log(crim), log(tax)))
p10 <- ggplot(data) +geom_point(aes(log(crim), log(ptratio)))
p11 <- ggplot(data) +geom_point(aes(log(crim), log(b)))
p12 <- ggplot(data) +geom_point(aes(log(crim), log(lstat)))
p13 <- ggplot(data) +geom_point(aes(log(crim), log(medv)))

gridExtra::grid.arrange(p1,p2, p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13, ncol =3)



#Dis variable seems to have strong relationship with crim
fit1 <- lm(log(crim) ~ log(dis), data)

rmse(fit1, data)

#Parameters of fitted models

coef(fit1)




```


Plot the residuals of the fitted model
```{r}

#PLOTTING RESIDUALS FOR ALL THE VARIABLES
a1 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (zn), y = resid))+
  geom_point() 

a2 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (indus), y = resid))+
  geom_point() 

a3 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (chas), y = resid))+
  geom_boxplot()

a4 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (nox), y = resid))+
  geom_point() 

a5 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = age, y = resid))+
  geom_point() 


a6 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (rad), y = resid))+
  geom_point() 

a7 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (tax), y = resid))+
  geom_point() 

a8 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (ptratio), y = resid))+
  geom_point() 

a9 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (b), y = resid))+
  geom_point() 

a10 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (lstat), y = resid))+
  geom_point() 


a11 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (medv), y = resid))+
  geom_point() 

a12 <- data %>%
  add_residuals(fit1) %>%
  ggplot(aes(x = (rm), y = resid))+
  geom_point() 


gridExtra::grid.arrange(a2, a4, a6,  a9, a10, a11,a12, ncol = 3, top = grid::textGrob("Potential Variables to be added to the model") )







```

Zn - No pattern Found
indus - weak negative relation
chas - No Pattern Found
nox - weak positive relation
age - No Pattern Found
rad - positive relation
tax - No pattern found
ptratio - No pattern found
b - non- linear relation
lstat - weak positive
medv -weak negative relation
rm - non linear relation


Fit a new model for predicting per capita crime rate by town, adding or removing variables based on the
residual plots
```{r}
#fitting variables Of observed patterns
fit2 <- lm(log(crim) ~ log(dis) + log(indus) + log(nox) + log(rad)+ log(b)+ log(lstat)
               + log(medv) + log(rm), data)


rmse(fit2, data)

#Trying to adjust model by removing non-linear patterned variables and making transformations

fit3 <- lm(log(crim) ~ (dis) + log(indus) + log(nox) + (rad)+ log(lstat)
               + log(medv) , data )

rmse(fit3, data)

#New model seems to work better 

#plotting against residuals for checking scatterness
data %>%
  add_residuals(fit3) %>%
 ggplot(aes(x = medv, y = resid))+
  geom_point()
#All the predictors seems to be showing good relation with the response variable

summary(fit3)

  

```
New model performs better than previous with lower rmse and high R- squared value. Scattered pattern is found when plotted against residuals. summary shows almost all the variables are significant with relatively lower p-values.
