---
title: " US Department of Education’s Civil Rights"
author: "Aishwarya"
date: "10/16/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We would like to investigate whether Black students receive a disproportionate number of in-school suspensions

```{r, warning = FALSE, echo=TRUE}
library(readr)
library(ggplot2)
library(dplyr)
school <- read_csv("C:/Users/Aishwarya/Desktop/NEU/Introduction to Data management/hw2/Data Files and Layouts/CRDC 2015-16 School Data.csv")

```

Create a new data.frame with the following columns:
• The total number of students enrolled at each school
• The number of Black students enrolled at each school
• The total number of students who received one or more in-school suspension (including non-disabled
students and disabled students served by IDEA)
• The number of Black students who received one or more in-school suspension (including non-disabled
students and disabled students served by IDEA)
• The proportion of Black students at each school among all students
• The proportion of students who received one or more in-school suspension who are Black among all
suspended students
Plot the proportion of Black students at each school (on the x-axis) versus the proportion of suspended
students who are Black (on the y-axis). Include a smoothing line on the plot.

```{r pressure, echo=TRUE, warning = FALSE}
#CHECKING FOR PRESCENCE OF RESERVE CODES
any(school$TOT_ENR_F < 0)

#REPLACING EACH REQUIRIED COLOUMN CONTAINING RESERVE CODES WITH NA
school$TOT_ENR_M<-   ifelse(school$TOT_ENR_M < 0, NA, school$TOT_ENR_M)
school$TOT_ENR_F<-   ifelse(school$TOT_ENR_F < 0, NA, school$TOT_ENR_F)


school$SCH_ENR_BL_F<-   ifelse(school$SCH_ENR_BL_F< 0, NA, school$SCH_ENR_BL_F)
school$SCH_ENR_BL_M<-   ifelse(school$SCH_ENR_BL_M< 0, NA, school$SCH_ENR_BL_M)


school$TOT_DISCWDIS_ISS_IDEA_F <-   ifelse(school$TOT_DISCWDIS_ISS_IDEA_F < 0, NA, school$TOT_DISCWDIS_ISS_IDEA_F)
school$TOT_DISCWDIS_ISS_IDEA_M<-   ifelse(school$TOT_DISCWDIS_ISS_IDEA_M < 0, NA, school$TOT_DISCWDIS_ISS_IDEA_M)
school$TOT_DISCWODIS_ISS_M <-   ifelse(school$TOT_DISCWODIS_ISS_M < 0, NA, school$TOT_DISCWODIS_ISS_M )
school$TOT_DISCWODIS_ISS_F <-   ifelse(school$TOT_DISCWODIS_ISS_F < 0, NA, school$TOT_DISCWODIS_ISS_F)


school$SCH_DISCWODIS_ISS_BL_M <-   ifelse(school$SCH_DISCWODIS_ISS_BL_M < 0, NA, school$SCH_DISCWODIS_ISS_BL_M)
school$SCH_DISCWODIS_ISS_BL_F <-   ifelse(school$SCH_DISCWODIS_ISS_BL_F < 0, NA, school$SCH_DISCWODIS_ISS_BL_F)
school$SCH_DISCWDIS_ISS_IDEA_BL_F <-   ifelse(school$SCH_DISCWDIS_ISS_IDEA_BL_F < 0, NA, school$SCH_DISCWDIS_ISS_IDEA_BL_F)
school$SCH_DISCWDIS_ISS_IDEA_BL_M <-   ifelse(school$SCH_DISCWDIS_ISS_IDEA_BL_M < 0, NA, school$SCH_DISCWDIS_ISS_IDEA_BL_M)


#CREATING NEW DATAFRAME AND INSERTING NEW COMPUTED VARIABLES
d1 <- data.frame(transmute(school,
      Total_students = TOT_ENR_M + TOT_ENR_F,
      
      Total_Black_students = SCH_ENR_BL_M + SCH_ENR_BL_F,
      
      Total_in_suspension = TOT_DISCWDIS_ISS_IDEA_M + TOT_DISCWDIS_ISS_IDEA_F +TOT_DISCWODIS_ISS_M + TOT_DISCWODIS_ISS_F,
      
      Total_Black_in_suspension = SCH_DISCWODIS_ISS_BL_M + SCH_DISCWODIS_ISS_BL_F + SCH_DISCWDIS_ISS_IDEA_BL_M + SCH_DISCWDIS_ISS_IDEA_BL_F ,
      
      Proportion_Black = Total_Black_students/Total_students,
      
      Proportion_Black_in_suspension = Total_Black_in_suspension /Total_in_suspension))


#PLOTTING USING GEOM_BIN2D
ggplot(d1) + geom_bin2d(mapping = aes(x =Proportion_Black, y = Proportion_Black_in_suspension)) + geom_smooth(mapping = aes(x =Proportion_Black, y = Proportion_Black_in_suspension))

```
OBSERVATION:-

The plot describes a positive relation between proportion of blacks and proprtion of suspended blacks but we can observe that at a certain point on x-axis there are HIGHER values of blacks in suspension which indicates an OVER REPRESENTATION of Blacks in school suspensions.


```{r}
#CALCULATING OVERALL PROPORTIONS
overall_prop_black <- mean(d1$Proportion_Black, na.rm = TRUE)
overall_prop_black_susp  <- mean(d1$Proportion_Black_in_suspension, na.rm = TRUE)

overall_prop_black
overall_prop_black_susp 
```

OBSERVATION:-
From calculations we can say that black students are OVER REPRESENTED.

We would like to investigate whether disabled students are more often disciplined with corporal punishment.
Create a new data.frame containing only schools that use corporal punishment with the following columns:
• The total number of students enrolled at each school
• The number of disabled students (served by IDEA) at each school
• The total number of students who were disciplined with corporal punishment
• The number of disabled students (served by IDEA) who were disciplined with corporal punishment
• The proportion of disabled students (served by IDEA) at each school among of all students
• The proportion of students who were disciplined with corporal punishment who are disabled (served by
IDEA) among all disciplined students.

Plot the proportion of disabled students at each school (on the x-axis) versus the proportion of disciplined
students who are disabled (on the y-axis).
```{r, warning = FALSE}

#REPLACING EACH REQUIRIED COLOUMN CONTAINING RESERVE CODES WITH NA
school$TOT_IDEAENR_M <-   ifelse(school$TOT_IDEAENR_M < 0, NA, school$TOT_IDEAENR_M)
school$TOT_IDEAENR_F <-   ifelse(school$TOT_IDEAENR_F < 0, NA, school$TOT_IDEAENR_F)


school$TOT_DISCWODIS_CORP_M <-   ifelse(school$TOT_DISCWODIS_CORP_M < 0, NA, school$TOT_DISCWODIS_CORP_M)
school$TOT_DISCWODIS_CORP_F <-   ifelse(school$TOT_DISCWODIS_CORP_F < 0, NA, school$TOT_DISCWODIS_CORP_F)
school$TOT_DISCWDIS_CORP_IDEA_M <-   ifelse(school$TOT_DISCWDIS_CORP_IDEA_M < 0, NA, school$TOT_DISCWDIS_CORP_IDEA_M)
school$TOT_DISCWDIS_CORP_IDEA_F <-   ifelse(school$TOT_DISCWDIS_CORP_IDEA_F < 0, NA, school$TOT_DISCWDIS_CORP_IDEA_F)

#CREATING NEW DATAFRAME AND INSERTING NEW COMPUTED VARIABLES
d2 <- data.frame(transmute(school,
      Total_students = TOT_ENR_M + TOT_ENR_F,
      
      Total_disable_students = TOT_IDEAENR_M + TOT_IDEAENR_F,
      
      Total_students_punishment =TOT_DISCWODIS_CORP_M+ TOT_DISCWODIS_CORP_F+TOT_DISCWDIS_CORP_IDEA_M+ TOT_DISCWDIS_CORP_IDEA_F,
      
      Disabled_punished = TOT_DISCWDIS_CORP_IDEA_M+ TOT_DISCWDIS_CORP_IDEA_F,
      
      Proportion_disabled = Total_disable_students/Total_students,
      
      Proportion_disabled_punished = Disabled_punished/ Total_students_punishment))


#PLOTTING USING GEOM_BIN2D
ggplot(d2) + geom_bin2d(mapping = aes(x =  Proportion_disabled, y = Proportion_disabled_punished )) + geom_smooth(mapping = aes(x =  Proportion_disabled, y = Proportion_disabled_punished ))
```

OBSERVATION:-

The relationship seems to be almost constant


```{r, warning = FALSE}
#CALCULATING OVERALL PROPORTIONS
overall_prop_disabled = mean(d2$Proportion_disabled, na.rm = TRUE)
overall_prop_disabled_punished = mean(d2$Proportion_disabled_punished, na.rm = TRUE)


overall_prop_disabled
overall_prop_disabled_punished

```
OBSERVATION:-

As said before, the variation slightly differs. so we can say that disabled students are equally treated with others when it comes to corporal punishment.

We would like to investigate whether Black and Hispanic students are over- or under-represented in Gifted &
Talented programs.
Create a new data.frame containing only schools with a Gifted & Talented program with the following
columns:
• The total number of students enrolled at each school
• The number of Black and Hispanic students at each school
• The total number of students in the school’s GT program
• The number of students in the GT program who are Black or Hispanic
• The proportion of students at each school who are Black or Hispanic among all students
• The proportion of students in the GT program who are Black or Hispanic among students in the GT
program
Plot the proportion of Black and Hispanic students at each school (on the x-axis) versus the proportion of
GT students who Black or Hispanic (on the y-axis).
```{r, warning = FALSE}

#REPLACING EACH REQUIRIED COLOUMN CONTAINING RESERVE CODES WITH NA
school$SCH_ENR_HI_M <-   ifelse(school$SCH_ENR_HI_M < 0, NA, school$SCH_ENR_HI_M)
school$SCH_ENR_HI_F <-   ifelse(school$SCH_ENR_HI_F < 0, NA, school$SCH_ENR_HI_F)


school$TOT_GTENR_M <-   ifelse(school$TOT_GTENR_M < 0, NA, school$TOT_GTENR_M)
school$TOT_GTENR_F <-   ifelse(school$TOT_GTENR_F < 0, NA, school$TOT_GTENR_F)

school$SCH_GTENR_BL_M <-   ifelse(school$SCH_GTENR_BL_M < 0, NA, school$SCH_GTENR_BL_M)
school$SCH_GTENR_BL_F <-   ifelse(school$SCH_GTENR_BL_F < 0, NA, school$SCH_GTENR_BL_F)

school$SCH_GTENR_HI_M <-   ifelse(school$SCH_GTENR_HI_M< 0, NA, school$SCH_GTENR_HI_M)
school$SCH_GTENR_HI_F <-   ifelse(school$SCH_GTENR_HI_F< 0, NA, school$SCH_GTENR_HI_F)


#CREATING NEW DATAFRAME AND INSERTING NEW COMPUTED VARIABLES
d3 <- data.frame(transmute(school,
      Total_students= TOT_ENR_M + TOT_ENR_F,
      Black_and_Hispanic = SCH_ENR_HI_M + SCH_ENR_HI_F + SCH_ENR_BL_M + SCH_ENR_BL_F,
      Total_students_GT = TOT_GTENR_M + TOT_GTENR_F,
      Black_or_Hispanic_GT= SCH_GTENR_BL_M + SCH_GTENR_BL_F+ SCH_GTENR_HI_M + SCH_GTENR_HI_F,
      Proportion_Black_or_Hispanic= Black_and_Hispanic/Total_students,
      Proportion_Black_or_Hispanic_GT= Black_or_Hispanic_GT/Total_students_GT))


#PLOTTING WITH GEOM_BIN2D
ggplot(d3) +
  geom_bin2d(mapping = aes(x =  Proportion_Black_or_Hispanic, y = Proportion_Black_or_Hispanic_GT)) +
  geom_smooth(mapping = aes(x =  Proportion_Black_or_Hispanic, y = Proportion_Black_or_Hispanic_GT))


```

OBSERVATION:-

Clearly, there is a positive relation but with negative slope between both proportions. At any point on Proportion of Black/Hispanic there exists corresponding point on Proportion of gifted Black/Hispanic with LOWER VALUE which is UNDER REPRESENTATION.

```{r, warning = FALSE}

#CALCULATING OVERALL PROPORTIONS
overall_prop_B_H= mean(d3$Proportion_Black_or_Hispanic, na.rm = TRUE)
overall_prop_B_H_GT = mean(d3$Proportion_Black_or_Hispanic_GT, na.rm = TRUE)

overall_prop_B_H
overall_prop_B_H_GT
```

OBSERVATION:-

As, 
overall proportion of Gifted Black and Hispanic is LOWER than overall proportion of Black and Hispanic, hence it is UNDER REPRESENTATION.
