---
title: "hw3-Aishwarya Vantipuli"
author: "Aishwarya"
date: "February 12, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

############################################## PART-A #########################################

## PROBLEM-1
```{r cars}
#Loading Libraries
library(dplyr)
library(ggplot2)
library(readr)
library(dplyr)






```

Filtering the data to include only the authors for whom a gender was predicted as ‘male’ or ‘female’ with a
probability of 0.95 or greater, and then create a bar plot showing the total number of distinct male and
female authors published each year

```{r , echo= TRUE, warning=FALSE}
#install.packages(c("DBI", "RSQLite", "dbplyr"))
#install.packages("RMySQL")


library("dbplyr")
library("RMySQL")
library("RSQLite")
library("DBI")



#cREATING cONNECTION
con <- dbConnect(MySQL(), 
                 user = 'root', 
                 password = 'Premkumar007', 
                 host = 'localhost',
                 dbname = 'dblp')
#LOADING TABLES
dblp_gen <- tbl(con, "general")

dblp_aut <- tbl(con, "authors")



total <-dblp_aut %>% left_join(dblp_gen) %>% group_by(year) %>%
  filter(gender %in% c("M", "F") & prob >= 0.95)   %>%
  summarise(Total_aut = n_distinct(name)) 

Total_sex <- dblp_aut %>% left_join(dblp_gen) %>% group_by(year, gender) %>%
  filter(gender %in% c("M", "F") & prob >= 0.95)   %>%
  summarise(Total_aut_sex = n_distinct(name)) 


#Prob 3
Total_sex %>%collect() %>%
  
  ggplot() +geom_col(aes(x = year,y = Total_aut_sex, fill = gender)) +
  
   ggtitle('Barplot of No. of authors in male and female within each Year')

  
#prob 5



  



```

There are no female authors untill arround 1970's. Female authors are increasing at lower rates comapared to male. There seems to be decrease in authors publications in recent years.


```{r , echo= TRUE, warning=FALSE}

#prob 4

right_join(Total_sex,total) %>%
  group_by(year, gender) %>% 
  summarise(prop = Total_aut_sex/Total_aut) %>% 
  ggplot() + geom_col(aes(x= year, y = prop, fill = gender)) +
  ggtitle("Proportion of male and female authors in each year")

```

Proportion of male is way higher than female and there are no female authors untill around 1965

##PROBLEM 5
```{r , echo=TRUE, warning=FALSE}
#prob 5


dblp_aut %>% left_join(dblp_gen) %>% group_by(year, gender, pos = 0) %>%
  filter(gender %in% c("M", "F") & prob > 0.95) %>% 
  summarise(pub_num = n_distinct(k)) %>% 
  ggplot() + geom_col(aes(x = gender,y = pub_num, fill = gender)) +
  ggtitle('No. of publications')
```
There are lower number of female publications than male
```{r , echo = TRUE}




```

