---
title: "Navajo Nation Water Quality"
author: "Aishwarya"
date: "January 31, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(dplyr)


#install.packages("maps")
#install.packages("mapproj")
#install.packages("measurements")



```

```{r}

#install.packages("maps")
#install.packages("mapproj")
#install.packages("measurements")
```

Mutate the dataset to replace the negative values of Radium-228 with 0, then filter the dataset to remove
any sites with “Unknown Risk” for the EPA risk rating.
Visualize the distribution of Radium-228 within each EPA section and each risk level

```{r, echo=TRUE, warning = FALSE}


data <- read_csv("C:/Users/Aishwarya/Desktop/NEU/Introduction to Data management/hw2/NavajoWater.csv" )

```

```{r, echo=TRUE, warning = FALSE}


#Removing Negative Values in Amount of Radium228 Coloumn
old_col <- data$`Amount of Radium228`
new_col <- ifelse(old_col < 0, 0, old_col)
data <- mutate(data,'Amount of Radium228'= new_col)

#Printing head
head(data$`Amount of Radium228`)

#Removing row containing 'Unknown Risk'
data <- data[-(which(data$`US EPA Risk Rating` == 'Unknown Risk')), ]

#Plotting Boxplots
ggplot(data = data, mapping = aes( x =`US EPA Risk Rating`, y =`Amount of Radium228`)) + geom_boxplot() + facet_wrap(~`Which EPA Section is This From?`)

```

OBSERVATION:-

Section 3,4,7 have comparatively more amount of radium228 which are of higher risk than other sections.



Create a map of the region (you may want to adjust the plotting limits to an appropriate “zoom” level) showing
the locations of the water sampling sites, along with the EPA risk and the concentration of Radium-228 for each location
```{r cars, echo=TRUE, warning = FALSE}

#PLOTING FOUR CORNERS
four_corners <- ggplot2::map_data("state", region = c("arizona", "New Mexico", "Utah", "Colorado"))

#CONVERTING LAT AND LONG DATA
data$Longitude <- measurements::conv_unit(data$Longitude, "deg_min_sec", "dec_deg")
data$Latitude <- measurements::conv_unit(data$Latitude, "deg_min_sec", "dec_deg")

#PLOTING POINTS ON FOUR CORNERS
ggplot(four_corners) +
  geom_polygon(mapping=aes(x=long,
                           y=lat,
                           group = group),
               fill=NA,
               color="black") +
 geom_point(data, mapping = aes(x = -as.double(Longitude), 
                                 y = as.double(Latitude),
                                 shape = `US EPA Risk Rating`,
                                 color = `Amount of Radium228`)) +
#ZOOMING  
  coord_fixed(xlim = c(-112.1, -109.1),  ylim = c(35.1,37.1), ratio = 1.3) +
  ggtitle("Locations of Water Sampling sites")



```
